---
title: "Simulate NeymanScott Point Processes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulate NeymanScott Point Processes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.asp=1,
  message=FALSE,
  warning=FALSE
)
```

Load the necessary libraries: 

```{r setup,results='hide'}
library(ggplot2);
library(spatstat);
library(rSimSpatialProcessSampling);
```

```{r runSims}
  #--simulate a Neyman-Scott point process 9 times

  #--create simulated clusters assuming a rotated 2d normal distribution for
  #----the dispersion within a cluster
  theta=45; sigx=0.05; sigy=0.01;
  res = spatstat.random::rNeymanScott(
                            kappa=4,
                            expand=3,
                            rcluster=clusterSims_GenerateNormalCluster,
                            n=1000,
                            sigx=sigx,
                            sigy=sigy,
                            theta=theta,
                            win=spatstat.geom::owin(c(0,1),c(0,1)),
                            lmax=NULL,
                            nsim=9,
                            drop=FALSE,
                            nonempty=TRUE,
                            saveparents=TRUE);
  #--create tibbles for the clusters and "parents"
  resp = clusterPPSims_CreateResultsTibbles(res);
  #--create plots
  ps = ppSims_MakePlots(resp,all_parents=TRUE);
```

The realized clusters are shown in Figure \@ref(fig:fig1), with the clusters colored by cluster parent id:

```{r fig1, fig.cap="The realized clusters.",echo=FALSE}
  print(ps[[1]]);
```
  
The realized cluster parents, combined across all simulations, are shown in Figure \@ref(fig:fig2):

```{r fig2, fig.cap="The realized clusters.",echo=FALSE}
  print(ps[[2]]);
```
  
Figure \@ref(fig:fig3) shows all the candidate cluster parents considered for the individual simulations. 
Because individual clusters may extend across the simulated area boundary (x-coordinate range: [0,1],
y-coordinate range: [0,1]), cluster parents need to be selected from a larger area that depends on 
the cluster size.
```{r fig3, fig.cap=cap,eval.after="fig.cap",echo=FALSE}
  cap =paste0('Cluster parents considered as candidates for the simulations. ',
              'Only clusters based on the the "observed" parents are incuded in the simulations.');
  print(ps[[3]]);
```

Now analyze the results: 

```{r analyzeSims}
  #--calculate stats from realizations and print results as a table
  tbl = ppSims_CalcStats(resp);
cap = "Statistics from the simulated processes. For a Poisson process, the mean ~ the variance."
  tbl |> kableExtra::kbl(booktabs=TRUE,digits=1,caption=cap,
                         col.names=c("type","mean number","variance","std. dev.")) |>
         kableExtra::kable_styling();
  
  #--quantify the patterns using 100 x 100 quadrats
  lst = list();
  for (i in 1:length(res)){
    qc = spatstat.geom::quadratcount(res[[i]],nx=100,ny=100);
    #--turn the results into a tibble
    lst[[i]] = quadratcount_AsTbl(qc) |> dplyr::filter(!is.na(value)) |> dplyr::mutate(sim=i);
  }
  tbl_qcs = dplyr::bind_rows(lst); rm(lst);
```

The patterns of abundance resulting from the simulated point processes are shown in \@ref(fig:fig4): 

```{r fig4, fig.cap="Abundance patterns based on the simulated point processes within the designated area."}
  #--plot the results (don't show the underlying clusters)
  ggplot(tbl_qcs |> dplyr::filter(!is.na(value)),aes(x=x,y=y,width=width,height=height,fill=value)) +
    geom_tile() +
    scale_fill_viridis_c(option="plasma",name="number") +
  #  geom_point(data=resp$clusters,mapping=aes(x=x,y=y,colour=id),inherit.aes=FALSE,shape=".") +
    facet_wrap(~sim) + 
    guides(colour="none") +
    wtsPlots::getStdTheme() + theme(axis.title=element_blank());
```
